Strength, mating success, and immune and nutritional costs in a population sample of US women and men: a registered report

### 2024-09-20 paper.qmd
Minor change of title and other edits.

---
title: "Strength, mating success, and immune and dietary costs in a population sample of US women and men: a registered report"

@@

Predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women. According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., a sex:strength interaction (the hypothesis does allow some positive or negative relationship between female strength and mating success, however, which is represented by the grey uncertainty band on the female slope).
:::

@@

```{r}
#| label: tbl-predictions
#| tbl-cap: "Predictions. A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant."

tribble(
              ~Outcome, ~`Grip Strength Coefficient`, ~`Grip Strength X Sex Coefficient`,
    "Partnered status",     "+",             "-",       
   "Lifetime partners",     "+",             "-",       
  "Past year partners",     "0",             "0",       
    "Age at first sex",     "0",             "0",       
            "Immunity",     "-",            "0",       
              "Energy",     "0",              "",       
             "Protein",     "+",              "",       
  ) |>    
  tt() |> 
  style_tt(j=c(1), align = 'l') 
```
## Power

@@

## Predictions and Results 
```{r}
#| label: tbl-predictionsresults
#| tbl-cap: "Predictions. A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; a zero (0) indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant. ✓: Supported in confirmatory study. ❌: Not supported in confirmatory study. *The posthoc results found a significant negative association between lean mass and immunity, and a positive association with energy."

tribble(
              ~Outcome, ~Prediction, ~Result, ~Prediction, ~Result,
    "Partnered status",     "+",       "✓",       "-",       "✓",
   "Lifetime partners",     "+",       "✓",       "-",       "❌",
  "Past year partners",     "0",       "✓",       "0",       "✓",
    "Age at first sex",     "0",       "✓",       "0",       "✓",
            "Immunity",     "-",       "❌*",      "0",       "✓",
              "Energy",     "0",       "❌*",        "",        "",
             "Protein",     "+",       "✓",        "",        ""
  ) |>      
  tt() |> 
  style_tt(j=c(1,3,5), align = 'l') |> 
  group_tt(j = list("Grip strength coefficient" = 2:3, "Grip strength X sex coefficient" = 4:5))
```

# Post Hoc Analyse

@@

Finally, we fit our grip strength models on the combined data (2011-2014). Results qualitatively correspond to those seen in the pilot study, which is not surprising because our models were developed on the pilot data. See @sfig-matingGH - @sfig-intakeGH.

@@

Our registered report had one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females. Our aim was not to test all possible byproduct hypotheses, nor was it possible to do so. We were unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS) that might influence the relationship between grip strength and our measures of mating success. Likewise, NHANES does not collect data on sociosexuality. As described in our methods, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from mating behavior for both men and women. Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993]. In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners (`{r} signif(100*sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)/nrow(designsH$d.design.adults$variables), 3)`%) on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis. Ideally, however, analyses would account for the full range of variation in sexual behavior. Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions. However, even our confirmatory results were based on generalized linear models, but there might be important non-linear relationships that would alter interpretations.

>>>>

---
title: "Strength, mating success, and immune and nutritional costs in a population sample of US women and men: a registered report"

@@

Theoretically predicted relationships between upper body strength and various indices of mating success, such as numbers of partners, partnered status, and age at first sex, for both men and women. According to the sexual selection hypothesis, sexual dimorphism in upper body strength is explained by a positive effect on male mating success that is greater than its effect on female mating success, i.e., a sex:strength interaction (the hypothesis does allow some positive or negative relationship between female strength and mating success, however, which is represented by the grey uncertainty band on the female slope).
:::

@@

```{r}
#| label: tbl-predictions
#| tbl-cap: "Predictions. A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; ns indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant."

tribble(
              ~Outcome, ~`Grip Strength Coefficient`, ~`Grip Strength X Sex Coefficient`,
    "Partnered status",     "+",             "-",       
   "Lifetime partners",     "+",             "-",       
  "Past year partners",     "ns",            "ns",       
    "Age at first sex",     "ns",            "ns",       
            "Immunity",     "-",             "ns",       
              "Energy",     "ns",             "",       
             "Protein",     "+",              "",       
  ) |>    
  tt() |> 
  style_tt(j=c(1), align = 'l') 
```
## Power

@@

## Predictions and Results 
```{r}
#| label: tbl-predictionsresults
#| tbl-cap: "Predictions. A plus sign (+) indicates a predicted significantly positive coefficient in all models of grip strength; a negative sign (-) indicates a predicted significantly negative coefficient in all such models; ns indicates a predicted non-significant coefficient in any such model; and a blank indicates not relevant. ✓: Supported in confirmatory study. ❌: Not supported in confirmatory study. *The posthoc results found a significant negative association between lean mass and immunity, and a positive association grip strength and lean mass with energy."

tribble(
              ~Outcome, ~Prediction, ~Supported, ~Prediction, ~Supported,
    "Partnered status",     "+",       "✓",       "-",       "✓",
   "Lifetime partners",     "+",       "✓",       "-",       "❌",
  "Past year partners",     "ns",      "✓",       "ns",       "✓",
    "Age at first sex",     "ns",      "✓",       "ns",       "✓",
            "Immunity",     "-",       "❌*",     "ns",       "✓",
              "Energy",     "ns",      "❌*",      "",        "",
             "Protein",     "+",       "✓",        "",        ""
  ) |>      
  tt() |> 
  style_tt(j=c(1,3,5), align = 'l') |> 
  group_tt(j = list("Grip strength coefficient" = 2:3, "Grip strength X sex coefficient" = 4:5))
```

# Post Hoc Analyses

@@

Finally, we fit our grip strength models on the combined data (2011-2014). Results qualitatively correspond to those in the pilot study, which is not surprising because our models were developed on the pilot data. See @sfig-matingGH - @sfig-intakeGH.

@@

Our registered report had one aim: to test the prediction from the sexual selection hypothesis that the effect of upper body strength on reproductive success is significantly positive for males and significantly less so for females. Our aim was not to test all possible byproduct hypotheses, nor was it possible to do so. We were unable to control for in-utero testosterone exposure, or other conditions such as Polycystic Ovarian Syndrome (PCOS) that might influence the relationship between grip strength and our measures of mating success. Likewise, NHANES does not collect data on sociosexuality. As described in our methods, in industrialized populations like the U.S., widespread access to contraceptives uncouples reproductive success from mating behavior for both men and women. Measures of mating success, including partnered status, number of sexual partners, and age at first sexual intercourse, are used as proxies as they are assumed to have been strongly correlated with reproductive success under ancestral conditions [@perusse1993]. In addition, in confirmatory analyses of mating outcomes (but not immune, energy, or protein outcomes) we removed `{r} sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)` participants with 100 or more lifetime sexual partners (`{r} signif(100*sum(designsH$d.design.adults$variables$sex_partners>=100, na.rm = T)/nrow(designsH$d.design.adults$variables), 3)`%) on the grounds that these might involve evolutionarily novel patterns of behavior that would not conform to predictions of the sexual selection hypothesis. Ideally, however, analyses would account for the full range of variation in sexual behavior (we fit our confirmatory models on the complete data, and all results were very similar to those we report). Finally, all pilot results were obtained after considerable exploratory analysis, with a high risk of overfitting the data, and therefore our confirmatory results were critical to testing our predictions. However, our confirmatory results were based on generalized linear models, but there might be important non-linear relationships that would alter interpretations.
